---
title: "BST 281 FP Module 1"
author: "Romy Li"
date: "2026-05-08"
output: html_document
---

### 1. Unpack the archive and organize the files

```{r}
untar("GSE140510_RAW.tar", exdir = "GSE140510_RAW")
list.files("GSE140510_RAW")
```

### 2. Build a sample metadata sheet

```{r}
sample_info <- data.frame(
  gsm = c(
    "GSM4173504","GSM4173505","GSM4173506",
    "GSM4173507","GSM4173508","GSM4173509",
    "GSM4173510","GSM4173511","GSM4173512",
    "GSM4173513","GSM4173514","GSM4173515"
  ),
  sample_id = c(
    "WT1","WT2","WT3",
    "Trem2_KO1","Trem2_KO2","Trem2_KO3",
    "WT_5XFAD1","WT_5XFAD2","WT_5XFAD3",
    "Trem2_KO_5XFAD1","Trem2_KO_5XFAD2","Trem2_KO_5XFAD3"
  ),
  condition = c(
    rep("WT", 3),
    rep("Trem2_KO", 3),
    rep("WT_5XFAD", 3),
    rep("Trem2_KO_5XFAD", 3)
  ),
  replicate = rep(1:3, 4),
  stringsAsFactors = FALSE
)

sample_info$trem2_status <- ifelse(grepl("Trem2_KO", sample_info$condition), "KO", "WT")
sample_info$ad_status <- ifelse(grepl("5XFAD", sample_info$condition), "5XFAD", "non_5XFAD")

sample_info
```

### 3. Read each sample separately into Seurat

```{r}
library(Seurat)
library(Matrix)

raw_dir <- "GSE140510_RAW"

read_one_sample <- function(gsm, sample_name) {
  all_files <- list.files(raw_dir, full.names = TRUE)
  
  mtx_file <- all_files[grepl(gsm, all_files) & grepl("matrix", all_files)]
  bc_file  <- all_files[grepl(gsm, all_files) & grepl("barcode", all_files)]
  feat_file <- all_files[grepl(gsm, all_files) & grepl("features|genes", all_files)]
  
  counts <- ReadMtx(
    mtx = mtx_file,
    cells = bc_file,
    features = feat_file,
    feature.column = 2
  )
  
  obj <- CreateSeuratObject(
    counts = counts,
    project = sample_name,
    min.cells = 3,
    min.features = 100
  )
  
  obj$gsm <- gsm
  obj$sample_id <- sample_name
  
  obj
}

objs <- vector("list", nrow(sample_info))
names(objs) <- sample_info$sample_id

for (i in seq_len(nrow(sample_info))) {
  objs[[i]] <- read_one_sample(sample_info$gsm[i], sample_info$sample_id[i])
  
  objs[[i]]$condition <- sample_info$condition[i]
  objs[[i]]$replicate <- sample_info$replicate[i]
  objs[[i]]$trem2_status <- sample_info$trem2_status[i]
  objs[[i]]$ad_status <- sample_info$ad_status[i]
}

# Inspect one object
objs[[1]]
head(colnames(objs[[1]]))
head(rownames(objs[[1]]))

# Check the dimensions of all objects to make sure there's no outliers
sapply(objs, dim)
```

### 4. Compute QC metrics before filtering

```{r}
# Check mitochondrial gene prefix
grep("^mt-", rownames(objs[[1]]), value = TRUE)[1:20]

# Add mitochondrial percentage
for (i in seq_along(objs)) {
  objs[[i]][["percent.mt"]] <- PercentageFeatureSet(objs[[i]], pattern = "^mt-")
}

# Build a combined QC dataframe
qc_df <- do.call(rbind, lapply(names(objs), function(s) {
  md <- objs[[s]]@meta.data
  md$sample <- s
  md$cell <- rownames(md)
  md
}))

head(qc_df)
table(qc_df$sample)
```

Make a summary table

```{r}
qc_summary <- do.call(rbind, lapply(names(objs), function(s) {
  x <- subset(qc_df, sample == s)
  data.frame(
    sample = s,
    n_cells = nrow(x),
    nFeature_median = median(x$nFeature_RNA),
    nFeature_q01 = quantile(x$nFeature_RNA, 0.01),
    nFeature_q05 = quantile(x$nFeature_RNA, 0.05),
    nFeature_q95 = quantile(x$nFeature_RNA, 0.95),
    nFeature_q99 = quantile(x$nFeature_RNA, 0.99),
    nCount_median = median(x$nCount_RNA),
    nCount_q95 = quantile(x$nCount_RNA, 0.95),
    nCount_q99 = quantile(x$nCount_RNA, 0.99),
    percent_mt_median = median(x$percent.mt),
    percent_mt_q95 = quantile(x$percent.mt, 0.95),
    percent_mt_q99 = quantile(x$percent.mt, 0.99)
  )
}))

qc_summary
```

The table shows whether any sample has unusually low complexity or much higher mitochondrial percentage. It also gives an idea of what sensible cutoffs might be.

For better visualization, let's make three combined plots across all samples

```{r}
library(ggplot2)

ggplot(qc_df, aes(x = sample, y = nFeature_RNA)) +
  geom_violin(scale = "width") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(qc_df, aes(x = sample, y = nCount_RNA)) +
  geom_violin(scale = "width") +
  scale_y_log10() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(qc_df, aes(x = sample, y = percent.mt)) +
  geom_violin(scale = "width") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

I also created two sets of scatter plots with `nCount_RNA` on the x-axis and either `nFeature_RNA` or `percent.mt` on the y-axis. This gives a quick visual check for low-count debris, possible doublets, and any strange sample-specific structure. Random sub-sampling of 1,000 cells from each sample is employed for efficiency.

```{r}
set.seed(1)

qc_scatter <- do.call(rbind, lapply(split(qc_df, qc_df$sample), function(df) {
  df[sample(seq_len(nrow(df)), min(1000, nrow(df))), ]
}))

ggplot(qc_scatter, aes(x = nCount_RNA, y = nFeature_RNA)) +
  geom_point(size = 0.3, alpha = 0.4) +
  facet_wrap(~sample, scales = "free") +
  theme_bw()

ggplot(qc_scatter, aes(x = nCount_RNA, y = percent.mt)) +
  geom_point(size = 0.3, alpha = 0.4) +
  facet_wrap(~sample, scales = "free") +
  theme_bw()
```

### 5. Doing a first-pass QC filter

Based on the above results, I'm choosing the following first-pass filtering criteria:

- `nFeature_RNA >= 400`: the 1st-percentile values are around 407 to 460 for most samples, so 400 is a sensible lower bound that removes obvious low-quality barcodes without being too aggressive.

- `nFeature_RNA <= 6000`: Most samples have their 99th percentile around 4,800 to 5,900 genes, and only a sparse tail extends beyond that. Cells above 6,000 are likely enriched for doublets or multiplets, so 6,000 is a good upper cutoff.

- `nCount_RNA <= 20000`: The nCount distributions are pretty similar across samples, and only a small number of cells sit in the extreme high-count tail. A 20,000 cutoff removes those obvious outliers without trimming too much of the valid distribution.

- `percent.mt <= 10`: Most nuclei have extremely low mitochondrial percentage, which is what we expect for snRNA-seq, but a few samples have broader upper tails. Using 10% removes the clearly bad high-mito cells while avoiding disproportionately heavy filtering in samples like `Trem2_KO1` and `WT_5XFAD2`.

```{r}
objs_qc <- lapply(objs, function(x) {
  subset(
    x,
    subset = nFeature_RNA >= 400 &
             nFeature_RNA <= 6000 &
             nCount_RNA <= 20000 &
             percent.mt <= 10
  )
})

# Compare cell counts before and after
data.frame(
  sample = names(objs),
  before = sapply(objs, ncol),
  after = sapply(objs_qc, ncol)
)
```

The criteria don't appear excessively or disproportionally harsh on any sample.

### 6. Save the pre-integration handoff objects

These objects are useful for downstream analysis of other group members.

```{r}
saveRDS(objs, "module1_objects_prefilter.rds")
saveRDS(objs_qc, "module1_objects_postfilter_initial.rds")
write.csv(sample_info, "sample_info.csv", row.names = FALSE)
```

### 7. Normalize each sample and prepare for integration

I used RPCA instead of CCA and reference-based integration instead of pairwise integration to speed up the process since I'm running on a local machine with limited RAM.

```{r}
library(Seurat)

# Clean up memory before integration
rm(objs, qc_df, qc_scatter, qc_summary, i, features)
gc()

# Unique cell names
objs_qc <- Map(
  f = function(obj, nm) RenameCells(obj, add.cell.id = nm),
  obj = objs_qc,
  nm = names(objs_qc)
)

gc()

# Preprocess each sample for RPCA integration
objs_qc <- lapply(objs_qc, function(x) {
  DefaultAssay(x) <- "RNA"
  x <- NormalizeData(x, verbose = FALSE)
  x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000, verbose = FALSE)
  x <- ScaleData(x, verbose = FALSE)
  x <- RunPCA(x, npcs = 30, verbose = FALSE)
  x
})

gc()

# Choose one reference sample per biological condition
ref.samples <- c("WT2", "Trem2_KO2", "WT_5XFAD1", "Trem2_KO_5XFAD2")
ref.idx <- match(ref.samples, names(objs_qc))

# Select integration features
features <- SelectIntegrationFeatures(
  object.list = objs_qc,
  nfeatures = 2000
)

# Faster anchor finding: RPCA + reference-based integration
anchors <- FindIntegrationAnchors(
  object.list = objs_qc,
  anchor.features = features,
  reduction = "rpca",
  dims = 1:30,
  reference = ref.idx
)

gc()

# Integrate datasets
obj_integrated <- IntegrateData(
  anchorset = anchors,
  dims = 1:30
)

# Free memory once integration is done
rm(anchors, features)
gc()

# Downstream dimensional reduction and clustering
DefaultAssay(obj_integrated) <- "integrated"

obj_integrated <- ScaleData(obj_integrated, verbose = FALSE)
obj_integrated <- RunPCA(obj_integrated, npcs = 30, verbose = FALSE)
obj_integrated <- RunUMAP(obj_integrated, dims = 1:30)
obj_integrated <- FindNeighbors(obj_integrated, dims = 1:30)
obj_integrated <- FindClusters(obj_integrated, resolution = 0.3)

# Save result
saveRDS(obj_integrated, "module1_integrated_rpca_reference.rds")
```

### 8. Broad Cell Type Annotation with the Integrated Dataset

```{r, eval=FALSE}
# Load checkpoint if necessary
library(Seurat)
obj_integrated <- readRDS("module1_integrated_rpca_reference.rds")
```

Check RNA assay state before running marker discovery

```{r}
DefaultAssay(obj_integrated) <- "RNA"

print(Layers(obj_integrated[["RNA"]]))
print(table(obj_integrated$seurat_clusters))

sample_cluster_tab <- table(obj_integrated$sample_id, obj_integrated$seurat_clusters)
dim(sample_cluster_tab)
sample_cluster_tab[1:5, 1:10]
```

Join the RNA layers, run cluster markers, and make a compact top-marker table for annotation

```{r}
obj_integrated <- JoinLayers(obj_integrated, assay = "RNA")

Layers(obj_integrated[["RNA"]])

Idents(obj_integrated) <- "seurat_clusters"

all_markers <- FindAllMarkers(
  obj_integrated,
  assay = "RNA",
  only.pos = TRUE,
  min.pct = 0.25,
  logfc.threshold = 0.25
)

top_markers <- do.call(
  rbind,
  lapply(split(all_markers, all_markers$cluster), function(df) {
    df <- df[order(df$avg_log2FC, decreasing = TRUE), ]
    head(df[, c("cluster", "gene", "avg_log2FC", "pct.1", "pct.2", "p_val_adj")], 10)
  })
)

top_markers
```

Cluster-level annotation table

```{r}
# 2) canonical marker panels for broad cell types
marker_sets <- list(
  Microglia = c("Tmem119", "P2ry12", "Cx3cr1", "Aif1", "Tyrobp", "Csf1r"),
  Astrocyte = c("Aqp4", "Gfap", "Aldh1l1", "Slc1a3", "Aldoc", "Clu"),
  Oligodendrocyte = c("Mog", "Mbp", "Mobp", "Plp1", "Mag", "Cldn11", "Sox10"),
  OPC = c("Pdgfra", "Cspg4", "Vcan", "Olig1", "Olig2"),
  Neuron = c("Snap25", "Syt1", "Rbfox3", "Tubb3", "Slc17a7", "Gad1", "Gad2"),
  Endothelial = c("Cldn5", "Pecam1", "Kdr", "Flt1", "Emcn"),
  Pericyte_VSMC = c("Rgs5", "Pdgfrb", "Kcnj8", "Abcc9", "Acta2", "Mcam"),
  Ependymal_Choroid = c("Foxj1", "Ttr", "Ccdc153", "Pifo")
)

# 3) average RNA expression by cluster
avg_expr <- AverageExpression(
  obj_integrated,
  assays = "RNA",
  slot = "data",
  group.by = "seurat_clusters",
  verbose = FALSE
)$RNA

avg_expr_z <- t(scale(t(avg_expr)))
avg_expr_z[is.na(avg_expr_z)] <- 0

# 4) score each cluster against each marker panel
score_mat <- sapply(names(marker_sets), function(ct) {
  genes <- intersect(marker_sets[[ct]], rownames(avg_expr_z))
  if (length(genes) == 0) {
    rep(NA_real_, ncol(avg_expr_z))
  } else {
    colMeans(avg_expr_z[genes, , drop = FALSE])
  }
})

# rebuild score_df with cleaned cluster names
score_df <- data.frame(
  cluster = sub("^g", "", colnames(avg_expr_z)),
  as.data.frame(score_mat),
  check.names = FALSE
)

score_cols <- names(marker_sets)

score_df$provisional_celltype <- score_cols[
  max.col(score_df[, score_cols, drop = FALSE], ties.method = "first")
]

score_df$max_score <- apply(
  score_df[, score_cols, drop = FALSE],
  1,
  max,
  na.rm = TRUE
)

score_df$provisional_celltype[
  is.na(score_df$max_score) | score_df$max_score < 0.25
] <- "Unknown"

cluster_key <- merge(
  data.frame(cluster = as.character(sort(unique(obj_integrated$seurat_clusters)))),
  score_df[, c("cluster", "provisional_celltype", "max_score")],
  by = "cluster",
  all.x = TRUE
)

cluster_key$top_markers <- unname(top3_markers[cluster_key$cluster])

cluster_key
```

We can see that the panel-scoring algorithm is working but only partially. There're still many `Ependymal_choroid` calls, some `astrocyte` calls, and some `unknown` clusters. Let's inspect canonical markers directly by cluster.

```{r, fig.width=12, fig.height=5}
canonical_markers <- c(
  "Tmem119","P2ry12","Cx3cr1","Aif1","Tyrobp","Csf1r",          # microglia
  "Aqp4","Gfap","Aldh1l1","Slc1a3","Aldoc","Clu",              # astrocyte
  "Mog","Mbp","Mobp","Plp1","Mag","Cldn11","Sox10",            # oligodendrocyte
  "Pdgfra","Cspg4","Vcan","Olig1","Olig2",                     # OPC
  "Snap25","Syt1","Rbfox3","Tubb3","Slc17a7","Gad1","Gad2",    # neuron
  "Cldn5","Pecam1","Kdr","Flt1","Emcn",                        # endothelial
  "Rgs5","Pdgfrb","Kcnj8","Abcc9","Acta2","Mcam",              # pericyte/VSMC
  "Foxj1","Ttr","Ccdc153","Pifo"                               # ependymal/choroid
)

canonical_markers <- canonical_markers[canonical_markers %in% rownames(obj_integrated)]

p_dot <- DotPlot(
  obj_integrated,
  features = canonical_markers,
  assay = "RNA",
  group.by = "seurat_clusters"
) + RotatedAxis()

p_dot

dot_data <- DotPlot(
  obj_integrated,
  features = canonical_markers,
  assay = "RNA",
  group.by = "seurat_clusters"
)$data

write.csv(dot_data, "canonical_marker_dotplot_data.csv", row.names = FALSE)

canonical_markers
```

We can make some high-confidence broad calls based on the dot plot:

- `Oligodendrocyte`: 0, 26, 28

- `OPC`: 14

- `Microglia`: 9

- `Endothelial`: 25

- `Pericyte/VSMC`: 27

- `Astrocyte`: 8, 21

Clusters 1, 12, 19, and 30 are still ambiguous and needs further inspection.
```{r}
ambig_clusters <- c("1", "12", "19", "30")

ambig_top <- do.call(
  rbind,
  lapply(ambig_clusters, function(cl) {
    df <- subset(all_markers, cluster == cl)
    df <- df[order(df$avg_log2FC, decreasing = TRUE), ]
    head(df[, c("cluster", "gene", "avg_log2FC", "pct.1", "pct.2", "p_val_adj")], 15)
  })
)

ambig_top
```

Seeing the marker of the ambiguous clusters help us determined their identities. Below is the reviewed cluster key:
```{r}
cluster_key_reviewed <- data.frame(
  cluster = as.character(0:30),
  broad_celltype = c(
    "Oligodendrocyte",   # 0
    "Fibroblast_Stromal",# 1
    "Neuron",            # 2
    "Neuron",            # 3
    "Neuron",            # 4
    "Neuron",            # 5
    "Neuron",            # 6
    "Neuron",            # 7
    "Astrocyte",         # 8
    "Microglia",         # 9
    "Neuron",            # 10
    "Neuron",            # 11
    "Low_quality",       # 12
    "Neuron",            # 13
    "OPC",               # 14
    "Neuron",            # 15
    "Neuron",            # 16
    "Neuron",            # 17
    "Neuron",            # 18
    "Fibroblast_Stromal",# 19
    "Neuron",            # 20
    "Astrocyte",         # 21
    "Neuron",            # 22
    "Neuron",            # 23
    "Neuron",            # 24
    "Endothelial",       # 25
    "Oligodendrocyte",   # 26
    "Pericyte_VSMC",     # 27
    "Oligodendrocyte",   # 28
    "Neuron",            # 29
    "Astrocyte"          # 30
  ),
  stringsAsFactors = FALSE
)

obj_integrated$cluster_id <- as.character(obj_integrated$seurat_clusters)
obj_integrated$broad_celltype <- cluster_key_reviewed$broad_celltype[
  match(obj_integrated$cluster_id, cluster_key_reviewed$cluster)
]

obj_integrated$is_glia <- obj_integrated$broad_celltype %in% c(
  "Microglia", "Astrocyte", "Oligodendrocyte", "OPC"
)

write.csv(cluster_key_reviewed, "cluster_annotation_key_reviewed.csv", row.names = FALSE)

table(obj_integrated$broad_celltype)
table(obj_integrated$cluster_id, obj_integrated$broad_celltype)
```

I'd exclude cluster 12 from downstream handoff as it contains mostly `Low_quality` calls. The other clusters appear to be in good shape.
### 9. Exporting files to handoff to teammates
```{r}
# -----------------------------
# 1) save reviewed annotation key + marker tables
# -----------------------------
write.csv(cluster_key_reviewed, "cluster_annotation_key_reviewed.csv", row.names = FALSE)
write.csv(all_markers, "cluster_markers_all.csv", row.names = FALSE)
write.csv(top_markers, "cluster_top10_markers.csv", row.names = FALSE)

# -----------------------------
# 2) create handoff object excluding low-quality cluster
# -----------------------------
obj_handoff <- subset(obj_integrated, subset = broad_celltype != "Low_quality")

# keep RNA as default for downstream DE convenience
DefaultAssay(obj_handoff) <- "RNA"

# save both full annotated object and cleaned handoff object
saveRDS(obj_integrated, "module1_integrated_annotated_full.rds")
saveRDS(obj_handoff, "module1_integrated_handoff_no_lowq.rds")

# -----------------------------
# 3) export cell-level metadata
# -----------------------------
meta_full <- obj_integrated[[]]
meta_full$cell_barcode <- rownames(meta_full)

if ("umap" %in% names(obj_integrated@reductions)) {
  umap_full <- as.data.frame(Embeddings(obj_integrated, "umap"))
  colnames(umap_full) <- c("UMAP_1", "UMAP_2")
  meta_full <- cbind(meta_full, umap_full[rownames(meta_full), , drop = FALSE])
}

write.csv(meta_full, "cell_metadata_annotated_full.csv", row.names = FALSE)

meta_handoff <- obj_handoff[[]]
meta_handoff$cell_barcode <- rownames(meta_handoff)

if ("umap" %in% names(obj_handoff@reductions)) {
  umap_handoff <- as.data.frame(Embeddings(obj_handoff, "umap"))
  colnames(umap_handoff) <- c("UMAP_1", "UMAP_2")
  meta_handoff <- cbind(meta_handoff, umap_handoff[rownames(meta_handoff), , drop = FALSE])
}

write.csv(meta_handoff, "cell_metadata_handoff_no_lowq.csv", row.names = FALSE)

# -----------------------------
# 4) sample-by-celltype and sample-by-cluster tables
# -----------------------------
# start from observed metadata rows only
meta_handoff <- obj_handoff[[]]
meta_handoff$cell_barcode <- rownames(meta_handoff)

# one clean sample-level mapping
sample_info_handoff <- unique(
  meta_handoff[, c("sample_id", "condition", "replicate", "gsm", "trem2_status", "ad_status")]
)
sample_info_handoff <- sample_info_handoff[order(sample_info_handoff$sample_id), ]

write.csv(sample_info_handoff, "sample_info_handoff.csv", row.names = FALSE)

# observed-only sample x celltype counts
sample_celltype_counts <- aggregate(
  cell_barcode ~ sample_id + condition + replicate + broad_celltype,
  data = meta_handoff,
  FUN = length
)
colnames(sample_celltype_counts)[colnames(sample_celltype_counts) == "cell_barcode"] <- "n_cells"

sample_totals <- aggregate(
  n_cells ~ sample_id,
  data = sample_celltype_counts,
  FUN = sum
)
colnames(sample_totals)[2] <- "sample_total"

sample_celltype_counts <- merge(sample_celltype_counts, sample_totals, by = "sample_id")
sample_celltype_counts$prop_within_sample <- sample_celltype_counts$n_cells / sample_celltype_counts$sample_total
sample_celltype_counts$sample_total <- NULL

write.csv(sample_celltype_counts, "sample_by_celltype_counts_long.csv", row.names = FALSE)

# observed-only sample x cluster counts
sample_cluster_counts <- aggregate(
  cell_barcode ~ sample_id + condition + replicate + cluster_id,
  data = meta_handoff,
  FUN = length
)
colnames(sample_cluster_counts)[colnames(sample_cluster_counts) == "cell_barcode"] <- "n_cells"

sample_cluster_totals <- aggregate(
  n_cells ~ sample_id,
  data = sample_cluster_counts,
  FUN = sum
)
colnames(sample_cluster_totals)[2] <- "sample_total"

sample_cluster_counts <- merge(sample_cluster_counts, sample_cluster_totals, by = "sample_id")
sample_cluster_counts$prop_within_sample <- sample_cluster_counts$n_cells / sample_cluster_counts$sample_total
sample_cluster_counts$sample_total <- NULL

write.csv(sample_cluster_counts, "sample_by_cluster_counts_long.csv", row.names = FALSE)

# wide versions from corrected long tables
sample_celltype_wide <- xtabs(n_cells ~ sample_id + broad_celltype, data = sample_celltype_counts)
write.csv(as.data.frame.matrix(sample_celltype_wide), "sample_by_celltype_counts_wide.csv")

sample_cluster_wide <- xtabs(n_cells ~ sample_id + cluster_id, data = sample_cluster_counts)
write.csv(as.data.frame.matrix(sample_cluster_wide), "sample_by_cluster_counts_wide.csv")

# quick checks
cat("Corrected dimensions:\n")
print(dim(sample_celltype_counts))
print(dim(sample_cluster_counts))

cat("\nFirst few corrected rows:\n")
print(head(sample_celltype_counts))

cat("\nSample info:\n")
print(sample_info_handoff)

# -----------------------------
# 5) export microglia and glia subsets
# -----------------------------
glia_types <- c("Microglia", "Astrocyte", "Oligodendrocyte", "OPC")

microglia_obj <- subset(obj_handoff, subset = broad_celltype == "Microglia")
glia_obj <- subset(obj_handoff, subset = broad_celltype %in% glia_types)

saveRDS(microglia_obj, "microglia_subset.rds")
saveRDS(glia_obj, "glia_subset.rds")

microglia_meta <- microglia_obj[[]]
microglia_meta$cell_barcode <- rownames(microglia_meta)
write.csv(microglia_meta, "microglia_metadata.csv", row.names = FALSE)
write.csv(data.frame(cell_barcode = rownames(microglia_meta)),
          "microglia_barcodes.csv", row.names = FALSE)

glia_meta <- glia_obj[[]]
glia_meta$cell_barcode <- rownames(glia_meta)
write.csv(glia_meta, "glia_metadata.csv", row.names = FALSE)
write.csv(data.frame(cell_barcode = rownames(glia_meta)),
          "glia_barcodes.csv", row.names = FALSE)

# optional per-glial-class barcode files
for (ct in glia_types) {
  cells_ct <- rownames(meta_handoff)[meta_handoff$broad_celltype == ct]
  write.csv(
    data.frame(cell_barcode = cells_ct),
    paste0(tolower(ct), "_barcodes.csv"),
    row.names = FALSE
  )
}

# -----------------------------
# 6) quick summaries to verify export
# -----------------------------
cat("Handoff broad cell types:\n")
print(table(obj_handoff$broad_celltype))

cat("\nMicroglia cells:", ncol(microglia_obj), "\n")
cat("Glia cells:", ncol(glia_obj), "\n")

cat("\nDimensions of exported tables:\n")
print(dim(sample_celltype_counts))
print(dim(sample_cluster_counts))

cat("\nFirst few rows of sample_by_celltype_counts_long:\n")
print(head(sample_celltype_counts))
```

